---
title: "Bayesian-networks"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Collaborators : Andrej Betik(456604), Petr Klanica(469223), Marcel Kubík(433719)

```{r include=FALSE}
# Packages installation
install_pckgs_if_not_installed <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
pckgs <- c("ggplot2", "gridExtra", "mice", "caret")
suppressWarnings(install_pckgs_if_not_installed(pckgs))
```

## 1. Introduction

1. Dataset : German Credit Data, Learning Algorithm : Bayesian Generalized Linear Model (Andrej Betik)
2. 
3.

## 2. Exploratory Analysis

### 1. German Credit Data

This dataset contains data about people attempting to take a credit by the bank. Each person represents a Good Risk or Bad Risk for bank.

Good Risk:
An assumption that customer who takes a credit will be able to pay invoices. The bank tends to give a credit to this customer

Bad Risk:
A loan that is unlikely to be repaid because of bad credit history, insufficient income, or some other reason.

The original dataset contains 1000 entries with 20 categorial/symbolic attributes prepared by Prof. Hofmann. The original dataset can be found at https://archive.ics.uci.edu/ml/datasets/Statlog+%28German+Credit+Data%29

Hoverever the original dataset is kind of messy so I got simplified and well documented version from kaggle site https://www.kaggle.com/uciml/german-credit

#### The selected atributes are:
1. Age (numeric)
2. Sex (text: male, female)
3. Job (numeric: 0 - unskilled and non-resident, 1 - unskilled and resident, 2 - skilled, 3 - highly skilled)
4. Housing (text: own, rent, or free)
5. Saving accounts (text - little, moderate, quite rich, rich)
6. Checking account (numeric, in DM - Deutsch Mark)
7. Credit amount (numeric, in DM)
8. Duration (numeric, in month)
9. Purpose (text: car, furniture/equipment, radio/TV, domestic appliances, repairs, education, business, vacation/others)
10. Risk(Target value- Goor or Bad Risk)

First look at the data

```{r}
#loading the dataset
german_credit_data <- read.csv("data/german_credit_data.csv")
# deleting the index column, the author was using that to uniquely identify every row
german_credit_data = german_credit_data[,-1]
# showing all the attributes and values it can take
str(german_credit_data)
```

Proportion of Good / Bad risk

```{r}
barplot(c(table(german_credit_data$Risk)), col = c("red", "blue"))
```

Attribute Distributions
```{r}
par(mfrow=c(3,2))
hist(german_credit_data$Age, xlab = "Age", main = "Age Distribution")
hist(german_credit_data$Job, xlab = "Job  (0-non-resident,1-resident,2-skilled,3-highly skilled)", main = "Job Distribution")
hist(german_credit_data$Credit.amount, xlab = "Credit Amount", main = "Credit Amount Distribution")
hist(german_credit_data$Duration, xlab = "Duration", main = "Duration Distribution")
hist(as.numeric(german_credit_data$Housing), xlab = "Housing  (1-free, 2-own, 3-rent)", main = "Housing Distribution")
barplot(c(table(german_credit_data$Sex)))

```


Inspection of interesting data features
```{r}
# People at very young age seem to have worse credit worthiness
p1 <- ggplot(german_credit_data, aes(x=Age, color=Risk)) + geom_density()
# The shorter the credit loan duration, the better the chance of getting loan"
p2 <- ggplot(german_credit_data, aes(x=Duration, color=Risk)) + geom_density()
# A trend of "Higher credit amount, better credit worthiness"
p3 <- ggplot(german_credit_data, aes(x=Credit.amount, color=Risk)) + geom_density()
grid.arrange(p1, p2, p3)
```

Inspection of NA values
```{r}
number_of_na_values <- sum(is.na(german_credit_data))
number_of_not_na_values <- sum(!is.na(german_credit_data))
# Proportion of number of NA values and number of normal(not NA values) out of all values
barplot(c(number_of_na_values, number_of_not_na_values), names.arg = c("NA", "Not NA"))
```

Na values occur only in attributes Savings Account and Checking account
```{r}
sum(is.na(german_credit_data))
sum(is.na(german_credit_data$Saving.accounts)) + sum(is.na(german_credit_data$Checking.account))

```

## 3. Preprocessing
### 1. German Credit Data
Since this dataset contains a lot of NA values, lets get rid of it with function mice
```{r}
# Number of NA values
sum(is.na(german_credit_data))
# First 5 NA values
head(german_credit_data[, c(5,6)])
german_credit_data <- complete(mice(german_credit_data))
sum(is.na(german_credit_data))
head(german_credit_data[, c(5,6)])
```

Apparantly this dataset is unballanced(has unequal number of instances for different classes)
```{r}
table(german_credit_data$Risk)
#TODO Over/Under Sampling

```


## 4. Learning Algorithms
### 1. Bayesian Generalized Linear Model
Tbe  generalized linear model (GLM) is a flexible generalization of ordinary linear regression that allows for response variables that have error distribution models other than a normal distribution. The GLM generalizes linear regression by allowing the linear model to be related to the response variable via a link function and by allowing the magnitude of the variance of each measurement to be a function of its predicted value.

The GLM consists of three elements:

1. An exponential family of probability distributions.
2. A linear predictor η = Xβ .
3. A link function g such that E(Y|X) = μ = g−1(η).

## 5. Evaluation
### 1. Bayesian Generalized Linear Model
```{r}
# TODO trainControl and cv method
set.seed(123)
indicies <- createDataPartition(german_credit_data$Risk, p = 0.7, list=FALSE)
german_credit_train <- german_credit_data[indicies, ]
german_credit_test <- german_credit_data[-indicies, ]
bayesglm_german_credit_model <- train(Risk ~ ., data = german_credit_train, method = "bayesglm")
bayesglm_german_credit_prediction <- predict(bayesglm_german_credit_model, german_credit_test)
confusionMatrix(bayesglm_german_credit_prediction, german_credit_test$Risk)

```

## 6. Conclusion
## 7. References
https://archive.ics.uci.edu/ml/datasets/Statlog+%28German+Credit+Data%29  
https://www.kaggle.com/uciml/german-credit  
https://en.wikipedia.org/wiki/Generalized_linear_model
